---
title: "IPEDS Acquisition"
author: "Sarah McDonald"
date: "October 19, 2019"
output: html_document
---
This document will discuss the process for acquiring five IPEDS data tables:

* Fall Enrollment - Race/ethnicity, gender, attendance status, and level of student

* Fall Enrollment - Major field of study, race/ethnicity, gender, attendance status, and level of student

* 12-Month Enrollment - 12-month unduplicated headcount: 2017-18

* Institutional Characteristics - Directory information

* Institutional Characteristics - Educational offerings, organization, services and athletic associations

# Set up
For the acquisition, four packages are needed: *gtools*, *dplyr*, *stringr*, and *stringi*. Additionally, we read in an csv file with the variables we want from each table.
```{r, warning= FALSE, message= FALSE}
library(gtools)
library(dplyr)
library(stringr) #to use str_trim
library(stringi)
variables <- read.csv("updated-variables.csv")
```

We create a function called "Upper_Converter" that changes all strings to uppercase. Then, we also create a function call get "IPEDSData" that allows acquisition of the IPEDS data from the IPEDS website. When the data is read in, all variable names are made uppercase, and a year column is added to each dataset.
```{r, warning = FALSE, message = FALSE}

Upper_Converter <- function(strings){
  toupper(str_trim(iconv(strings, "ASCII", "UTF-8", sub="")))
}

getIPEDSData <- function(year, survey_file, capital_survey_file, extra = "", capital_extra = ""){
  dataset <- NULL
  temp <- tempfile()
  download.file(paste("https://nces.ed.gov/ipeds/datacenter/data/",  capital_survey_file, year, capital_extra, ".zip", sep = ""), temp)
  dataset <- read.csv(unz(temp, paste(survey_file, year, extra, ".csv", sep = "")))
  unlink(temp)
  names(dataset) <- toupper(names(dataset))
  dataset <- dataset %>% mutate_if(is.factor, Upper_Converter)
  dataset$YEAR <- year
  dataset
}
```

# Institutional Characteristics - Directory information

We use the loops below to acquire directory information from IPEDS for 1995 - 2017. Directory information is available from 1986. We use different loops below as naming conventions for the various tables change prior to 2002. We will need to come up with CBSA and Lat/Long for missing values in earlier years
```{r, warning = FALSE, message = FALSE}
# 2002 - 2017
survey_file <- "hd"
capital_survey_file <- "HD"
extra <- ""
capital_extra <- ""

for (year in 2002:2017){
  assign(paste(survey_file, year, extra, sep = ""), 
    getIPEDSData(year, survey_file, capital_survey_file, extra, capital_extra))
}

# 2000 - 2001
survey_file <- "fa"
capital_survey_file <- "FA"
extra <- "hd"
capital_extra <- "HD"

for (year in 2000:2001){
  assign(paste(survey_file, year, extra, sep = ""), 
    getIPEDSData(year, survey_file, capital_survey_file, extra, capital_extra))
}

# 1999
survey_file <- "ic"
capital_survey_file <- "IC"
extra <- "_hd"
capital_extra <- "_HD"

for (year in 99){
  assign(paste(survey_file, year, extra, sep = ""), 
    getIPEDSData(year, survey_file, capital_survey_file, extra, capital_extra))
}
ic99_hd$YEAR <- "1999"

# 1998
survey_file <- "ic"
capital_survey_file <- "IC"
extra <- "hdac"
capital_extra <- "HDAC"

for (year in 98){
  assign(paste(survey_file, year, extra, sep = ""), 
    getIPEDSData(year, survey_file, capital_survey_file, extra, capital_extra))
}
ic98hdac$YEAR <- "1998"

# 1997
survey_file <- "ic"
capital_survey_file <- "ic"
extra <- "_hdr"
capital_extra <- "_HDR"

for (year in 9798){
  assign(paste(survey_file, year, extra, sep = ""), 
    getIPEDSData(year, survey_file, capital_survey_file, extra, capital_extra))
}
ic9798_hdr$YEAR <- "1997"

# 1995 - 1996
survey_file <- "ic"
capital_survey_file <- "ic"
extra <- "_a"
capital_extra <- "_A"

for (year in c(9596,9697)){
  assign(paste(survey_file, year, extra, sep = ""), 
    getIPEDSData(year, survey_file, capital_survey_file, extra, capital_extra))
}
ic9596_a$YEAR <- "1995"
ic9697_a$YEAR <- "1996"

survey_file <- "ic"
capital_survey_file <- "ic"
extra <- "_a"
capital_extra <- "_A"

for (year in c(1992:1994)){
   assign(paste(survey_file, year, extra, sep = ""), 
    getIPEDSData(year, survey_file, capital_survey_file, extra, capital_extra))
}

ic1992_a$YEAR <- "1992"
ic1993_a$YEAR <- "1993"
ic1994_a$YEAR <- "1994"

survey_file <- "ic"
capital_survey_file <- "ic"
extra <- "_hdr"
capital_extra <- "_HDR"

for (year in c(1991)){
   assign(paste(survey_file, year, extra, sep = ""), 
    getIPEDSData(year, survey_file, capital_survey_file, extra, capital_extra))
}

ic1991_hdr$YEAR <- "1991"

IC90HD

survey_file <- "ic"
capital_survey_file <- "IC"
extra <- "hd"
capital_extra <- "HD"

for (year in c(90)){
   assign(paste(survey_file, year, extra, sep = ""), 
    getIPEDSData(year, survey_file, capital_survey_file, extra, capital_extra))
}

ic90hd$YEAR <- "1990"


```

We then use the following code to select our variables of interest from each directory information table and merge the tables for each year into one large table.
```{r, warning = FALSE, message=FALSE}
HD <- mget(ls(pattern = c("(hd\\d{4})|(ic\\d+(hd\\w?\\w?|_\\w+))|(fa.+)")))

levels(variables$Variable.Name) <- c(levels(variables$Variable.Name),"YEAR") 
variables[nrow(variables) + 1,] = list(Table= factor("HD"),Variable.Number= 9999, Variable.Name= factor("YEAR"))
HD_var <- variables[variables$Table == "HD", 3]

HD <- lapply(HD, function(x) x[(names(x)) %in% HD_var])
HD_df <- do.call("smartbind", HD)
```

## 12-month unduplicated headcount

We use the loops below to acquire the 12-month unduplicated headcount information from IPEDS for 2001 - 2017. In 2008, the variable names changed from the old system to a new system, so I choose the new naming system. The old and new variables are included in 2008-2010. We changed the variable names in the tables prior to 2007 to the corresponding variable name that matches the new system. We do not collect data prior to 2001, as data is divided by undergraduates and graduates rather than race and gender. It is important to note that the 2001 data does not include the "EFFYLEV" variable.


```{r, warning = FALSE, message = FALSE}
survey_file <- "effy"
capital_survey_file <- "EFFY"
extra <- ""
capital_extra <- ""


for (year in 2002:2017) {
  assign(paste(survey_file, year, extra, sep = ""), 
         getIPEDSData(year, survey_file, capital_survey_file, extra, capital_extra))
}

colnames(effy2002)[colnames(effy2002)=="FYRACE21"] <- "EFYHISPT"
colnames(effy2003)[colnames(effy2003)=="FYRACE21"] <- "EFYHISPT"
colnames(effy2004)[colnames(effy2004)=="FYRACE21"] <- "EFYHISPT"
colnames(effy2005)[colnames(effy2005)=="FYRACE21"] <- "EFYHISPT"
colnames(effy2006)[colnames(effy2006)=="FYRACE21"] <- "EFYHISPT"
colnames(effy2007)[colnames(effy2007)=="FYRACE21"] <- "EFYHISPT"

colnames(effy2002)[colnames(effy2002)=="FYRACE24"] <- "EFYTOTLT"
colnames(effy2003)[colnames(effy2003)=="FYRACE24"] <- "EFYTOTLT"
colnames(effy2004)[colnames(effy2004)=="FYRACE24"] <- "EFYTOTLT"
colnames(effy2005)[colnames(effy2005)=="FYRACE24"] <- "EFYTOTLT"
colnames(effy2006)[colnames(effy2006)=="FYRACE24"] <- "EFYTOTLT"
colnames(effy2007)[colnames(effy2007)=="FYRACE24"] <- "EFYTOTLT"

colnames(effy2002)[colnames(effy2002)=="FYRACE15"] <- "EFYTOTLM"
colnames(effy2003)[colnames(effy2003)=="FYRACE15"] <- "EFYTOTLM"
colnames(effy2004)[colnames(effy2004)=="FYRACE15"] <- "EFYTOTLM"
colnames(effy2005)[colnames(effy2005)=="FYRACE15"] <- "EFYTOTLM"
colnames(effy2006)[colnames(effy2006)=="FYRACE15"] <- "EFYTOTLM"
colnames(effy2007)[colnames(effy2007)=="FYRACE15"] <- "EFYTOTLM"


#2008 has two counts, old and new, the number of missing values is the same for both, so I am choosing the new
length(is.na(effy2008$FYRACE21))
length(is.na(effy2008$EFYHISPT))

length(is.na(effy2009$FYRACE21))
length(is.na(effy2009$EFYHISPT))

length(is.na(effy2010$FYRACE21))
length(is.na(effy2010$EFYHISPT))

survey_file <- "ef"
capital_survey_file <- "EF"
extra <- "d1"
capital_extra <- "D1"

for (year in 2001){
  try(
    assign(paste(survey_file, year, extra, sep = ""), 
           getIPEDSData(year, survey_file, capital_survey_file, extra, capital_extra)))
}

#No "EFFYLEV" value

#Grand Total
colnames(ef2001d1)[colnames(ef2001d1)=="FYRACE17"] <- "EFYTOTLT"
 
#Total Men
colnames(ef2001d1)[colnames(ef2001d1)=="FYRACE15"] <- "EFYTOTLM"

#total Hispanic men
colnames(ef2001d1)[colnames(ef2001d1)=="FYRACE09"] <- "EFYHISPM"
colnames(ef2001d1)[colnames(ef2001d1)=="FYRACE10"] <- "EFYHISPW"

#total hispanic men and women
ef2001d1$EFYHISPT <- ef2001d1$EFYHISPW + ef2001d1$EFYHISPM

```

We then use the following code to select our variables of interest from each 12-month enrollment table and merge the tables for each year into one large table. 

```{r, warning = FALSE, message = FALSE}
# List of Data frames
EFFY <- mget(ls(pattern = "effy|^ef\\d{4}d\\d?$"))

levels(variables$Variable.Name) <- c(levels(variables$Variable.Name),"YEAR") 
variables[nrow(variables) + 1,] = list(Table= factor("EFFY"),Variable.Number= 9999, Variable.Name= factor("YEAR"))
EFFY_var <- variables[variables$Table == "EFFY", 3]

# Merge and and only use the columns in our variable list

EFFY <- lapply(EFFY, function(x) x[(names(x)) %in% EFFY_var])
EFFY_df <- do.call("smartbind", EFFY)

EFFY_df$EFFYLEV[EFFY_df$EFFYLEV == "1"] <- "Total"
EFFY_df$EFFYLEV[EFFY_df$EFFYLEV == "2"] <- "Undergraduates"
EFFY_df$EFFYLEV[EFFY_df$EFFYLEV == "4"] <- "Graduate"

EFFY_df$LSTUDY[EFFY_df$LSTUDY == "1"] <- "Undergraduates"
EFFY_df$LSTUDY[EFFY_df$LSTUDY == "2"] <- "First-Professional"
EFFY_df$LSTUDY[EFFY_df$LSTUDY == "3"] <- "Graduate"
EFFY_df$LSTUDY[EFFY_df$LSTUDY == "999"] <- "Total"

EFFY_df <- EFFY_df[, -7]
EFFY_df$LSTUDY[EFFY_df$LSTUDY == "First-Professional"] <- "Graduate"
EFFY_df <- EFFY_df %>% group_by(UNITID, LSTUDY, YEAR) %>% summarize(EFYTOTLM = sum(EFYTOTLM, na.rm = TRUE), EFYTOTLT = sum(EFYTOTLT, na.rm = TRUE), EFYHISPT =sum(EFYHISPT, na.rm = TRUE))


```

# Fall Enrollment - Race/ethnicity, gender, attendance status, and level of student
We use the loops below to acquire the fall enrollment data (Race/Ethnicty/Gender) from IPEDS for 2000 - 2017. In 2008, the variable names changed from the old system to a new system, so I choose the new naming system. We changed the variable names in the tables prior to 2008 to the corresponding variable name that matches the new system. *I have code that can acquire 1980, 1986-1999 if needed.*

```{r, warning = FALSE, message = FALSE}
survey_file <- "ef"
capital_survey_file <- "EF"
extra <- "a"
capital_extra <- "A"

for (year in 2000:2017){
  assign(paste(survey_file, year, extra, sep = ""), 
         getIPEDSData(year, survey_file, capital_survey_file, extra, capital_extra))
}

#fix 2000
#2000 total men
colnames(ef2000a)[colnames(ef2000a)=="EFRACE15"] <- "EFTOTLM"
#2000 total women
colnames(ef2000a)[colnames(ef2000a)=="EFRACE16"] <- "EFTOTLW"
#total men and women
ef2000a$EFTOTLT <- ef2000a$EFTOTLM + ef2000a$EFTOTLW


#2000 total Hispanic men
colnames(ef2000a)[colnames(ef2000a)=="EFRACE09"] <- "EFHISPM"
colnames(ef2000a)[colnames(ef2000a)=="EFRACE10"] <- "EFHISPW"

#total hispanic men and women
ef2000a$EFHISPT <- ef2000a$EFHISPW + ef2000a$EFHISPM


#fix 2001
#2001 total men
colnames(ef2001a)[colnames(ef2001a)=="EFRACE15"] <- "EFTOTLM"
#2001 total women
colnames(ef2001a)[colnames(ef2001a)=="EFRACE16"] <- "EFTOTLW"
#total men and women
ef2001a$EFTOTLT <- ef2001a$EFTOTLM + ef2001a$EFTOTLW

#2001 total Hispanic 
colnames(ef2001a)[colnames(ef2001a)=="EFRACE09"] <- "EFHISPM"
colnames(ef2001a)[colnames(ef2001a)=="EFRACE10"] <- "EFHISPW"

#total hispanic men and women
ef2001a$EFHISPT <- ef2001a$EFHISPW + ef2001a$EFHISPM

#fix 2002
#grand total
colnames(ef2002a)[colnames(ef2002a)=="EFRACE24"] <- "EFTOTLT"
#total men
colnames(ef2002a)[colnames(ef2002a)=="EFRACE15"] <- "EFTOTLM"
# total hispanic men and women
colnames(ef2002a)[colnames(ef2002a)=="EFRACE21"] <- "EFHISPT"

#fix 2003
#grand total
colnames(ef2003a)[colnames(ef2003a)=="EFRACE24"] <- "EFTOTLT"
#total men
colnames(ef2003a)[colnames(ef2003a)=="EFRACE15"] <- "EFTOTLM"
# total hispanic men and women
colnames(ef2003a)[colnames(ef2003a)=="EFRACE21"] <- "EFHISPT"

#fix 2004
#grand total
colnames(ef2004a)[colnames(ef2004a)=="EFRACE24"] <- "EFTOTLT"
#total men
colnames(ef2004a)[colnames(ef2004a)=="EFRACE15"] <- "EFTOTLM"
# total hispanic men and women
colnames(ef2004a)[colnames(ef2004a)=="EFRACE21"] <- "EFHISPT"

#fix 2005
#grand total
colnames(ef2005a)[colnames(ef2005a)=="EFRACE24"] <- "EFTOTLT"
#total men
colnames(ef2005a)[colnames(ef2005a)=="EFRACE15"] <- "EFTOTLM"
# total hispanic men and women
colnames(ef2005a)[colnames(ef2005a)=="EFRACE21"] <- "EFHISPT"

#fix 2006
#grand total
colnames(ef2006a)[colnames(ef2006a)=="EFRACE24"] <- "EFTOTLT"
#total men
colnames(ef2006a)[colnames(ef2006a)=="EFRACE15"] <- "EFTOTLM"
# total hispanic men and women
colnames(ef2006a)[colnames(ef2006a)=="EFRACE21"] <- "EFHISPT"

#fix 2007
#grand total
colnames(ef2007a)[colnames(ef2007a)=="EFRACE24"] <- "EFTOTLT"
#total men
colnames(ef2007a)[colnames(ef2007a)=="EFRACE15"] <- "EFTOTLM"
# total hispanic men and women
colnames(ef2007a)[colnames(ef2007a)=="EFRACE21"] <- "EFHISPT"


survey_file <- "ef"
capital_survey_file <- "EF"
extra <- "_anr"
capital_extra <- "_ANR"

for (year in 95:99){
  assign(paste(survey_file, year, extra, sep = ""), 
         getIPEDSData(year, survey_file, capital_survey_file, extra, capital_extra))
}

for (year in 1994){ 
 assign(paste(survey_file, year, extra, sep = ""), 
         getIPEDSData(year, survey_file, capital_survey_file, extra, capital_extra))
}

ef1994_anr$year <- "1994"
ef95_anr$year <- "1995"
ef96_anr$year <- "1996"
ef97_anr$year <- "1997"
ef98_anr$year <- "1998"
ef99_anr$year <- "1999"

colnames(ef99_anr)[colnames(ef99_anr)=="EFRACE24"] <- "EFTOTLT"
colnames(ef99_anr)[colnames(ef99_anr)=="EFRACE15"] <- "EFTOTLM"
colnames(ef99_anr)[colnames(ef99_anr)=="EFRACE21"] <- "EFHISPT"

colnames(ef99_anr)[colnames(ef99_anr)=="EFRACE24"] <- "EFTOTLT"
colnames(ef99_anr)[colnames(ef99_anr)=="EFRACE15"] <- "EFTOTLM"
colnames(ef99_anr)[colnames(ef99_anr)=="EFRACE21"] <- "EFHISPT"

colnames(ef98_anr)[colnames(ef98_anr)=="EFRACE15"] <- "EFTOTLM"
colnames(ef98_anr)[colnames(ef98_anr)=="EFRACE16"] <- "EFTOTLW"
ef98_anr$EFTOTLT <- ef98_anr$EFTOTLM + ef98_anr$EFTOTLW
colnames(ef98_anr)[colnames(ef98_anr)=="EFRACE09"] <- "EFHISPM"
colnames(ef98_anr)[colnames(ef98_anr)=="EFRACE10"] <- "EFHISPW"
ef98_anr$EFHISPT <- ef98_anr$EFHISPW + ef98_anr$EFHISPM

colnames(ef97_anr)[colnames(ef97_anr)=="EFRACE15"] <- "EFTOTLM"
colnames(ef97_anr)[colnames(ef97_anr)=="EFRACE16"] <- "EFTOTLW"
ef97_anr$EFTOTLT <- ef97_anr$EFTOTLM + ef97_anr$EFTOTLW
colnames(ef97_anr)[colnames(ef97_anr)=="EFRACE09"] <- "EFHISPM"
colnames(ef97_anr)[colnames(ef97_anr)=="EFRACE10"] <- "EFHISPW"
ef97_anr$EFHISPT <- ef97_anr$EFHISPW + ef97_anr$EFHISPM

colnames(ef96_anr)[colnames(ef96_anr)=="EFRACE15"] <- "EFTOTLM"
colnames(ef96_anr)[colnames(ef96_anr)=="EFRACE16"] <- "EFTOTLW"
ef96_anr$EFTOTLT <- ef96_anr$EFTOTLM + ef96_anr$EFTOTLW
colnames(ef96_anr)[colnames(ef96_anr)=="EFRACE09"] <- "EFHISPM"
colnames(ef96_anr)[colnames(ef96_anr)=="EFRACE10"] <- "EFHISPW"
ef96_anr$EFHISPT <- ef96_anr$EFHISPW + ef96_anr$EFHISPM

colnames(ef95_anr)[colnames(ef95_anr)=="EFRACE15"] <- "EFTOTLM"
colnames(ef95_anr)[colnames(ef95_anr)=="EFRACE16"] <- "EFTOTLW"
ef95_anr$EFTOTLT <- ef95_anr$EFTOTLM + ef95_anr$EFTOTLW
colnames(ef95_anr)[colnames(ef95_anr)=="EFRACE09"] <- "EFHISPM"
colnames(ef95_anr)[colnames(ef95_anr)=="EFRACE10"] <- "EFHISPW"
ef95_anr$EFHISPT <- ef95_anr$EFHISPW + ef95_anr$EFHISPM

colnames(ef1994_anr)[colnames(ef1994_anr)=="EFRACE15"] <- "EFTOTLM"
colnames(ef1994_anr)[colnames(ef1994_anr)=="EFRACE16"] <- "EFTOTLW"
ef1994_anr$EFTOTLT <- ef1994_anr$EFTOTLM + ef1994_anr$EFTOTLW
colnames(ef1994_anr)[colnames(ef1994_anr)=="EFRACE09"] <- "EFHISPM"
colnames(ef1994_anr)[colnames(ef1994_anr)=="EFRACE10"] <- "EFHISPW"
ef1994_anr$EFHISPT <- ef1994_anr$EFHISPW + ef1994_anr$EFHISPM

survey_file <- "ef"
capital_survey_file <- "EF"
extra <- "_a"
capital_extra <- "_A"

for (year in 1991:1993){try(
    assign(paste(survey_file, year, extra, sep = ""), 
           getIPEDSData(year, survey_file, capital_survey_file, extra, capital_extra)))
}

#this gets 1990
for (year in 90){
    assign(paste(survey_file, year, extra, sep = ""), 
           getIPEDSData(year, survey_file, capital_survey_file, extra, capital_extra))
}

ef90_a$year <- "1990"
ef1991_a$year <- "1991"
ef1992_a$year <- "1992"
ef1993_a$year <- "1993"

colnames(ef1993_a)[colnames(ef1993_a)=="EFRACE15"] <- "EFTOTLM"
colnames(ef1993_a)[colnames(ef1993_a)=="EFRACE16"] <- "EFTOTLW"
ef1993_a$EFTOTLT <- ef1993_a$EFTOTLM + ef1993_a$EFTOTLW
colnames(ef1993_a)[colnames(ef1993_a)=="EFRACE09"] <- "EFHISPM"
colnames(ef1993_a)[colnames(ef1993_a)=="EFRACE10"] <- "EFHISPW"
ef1993_a$EFHISPT <- ef1993_a$EFHISPW + ef1993_a$EFHISPM

colnames(ef1992_a)[colnames(ef1992_a)=="EFRACE15"] <- "EFTOTLM"
colnames(ef1992_a)[colnames(ef1992_a)=="EFRACE16"] <- "EFTOTLW"
ef1992_a$EFTOTLT <- ef1992_a$EFTOTLM + ef1992_a$EFTOTLW
colnames(ef1992_a)[colnames(ef1992_a)=="EFRACE09"] <- "EFHISPM"
colnames(ef1992_a)[colnames(ef1992_a)=="EFRACE10"] <- "EFHISPW"
ef1992_a$EFHISPT <- ef1992_a$EFHISPW + ef1992_a$EFHISPM

colnames(ef1991_a)[colnames(ef1991_a)=="EFRACE15"] <- "EFTOTLM"
colnames(ef1991_a)[colnames(ef1991_a)=="EFRACE16"] <- "EFTOTLW"
ef1991_a$EFTOTLT <- ef1991_a$EFTOTLM + ef1991_a$EFTOTLW
colnames(ef1991_a)[colnames(ef1991_a)=="EFRACE09"] <- "EFHISPM"
colnames(ef1991_a)[colnames(ef1991_a)=="EFRACE10"] <- "EFHISPW"
ef1991_a$EFHISPT <- ef1991_a$EFHISPW + ef1991_a$EFHISPM

colnames(ef90_a)[colnames(ef90_a)=="EFRACE15"] <- "EFTOTLM"
colnames(ef90_a)[colnames(ef90_a)=="EFRACE16"] <- "EFTOTLW"
ef90_a$EFTOTLT <- ef90_a$EFTOTLM + ef90_a$EFTOTLW
colnames(ef90_a)[colnames(ef90_a)=="EFRACE09"] <- "EFHISPM"
colnames(ef90_a)[colnames(ef90_a)=="EFRACE10"] <- "EFHISPW"
ef90_a$EFHISPT <- ef90_a$EFHISPW + ef90_a$EFHISPM



```

We then use the following code to select our variables of interest from each fall enrollment(race/ethnicity/gender) table and merge the tables for each year into one large table. 

```{r, warning = FALSE, message = FALSE}
#2000-2017 (file is too big)
EFA <- mget(ls(pattern = c("ef\\d*(a|_)a?\\w?\\w?")))

levels(variables$Variable.Name) <- c(levels(variables$Variable.Name),"YEAR") 
variables[nrow(variables) + 1,] = list(Table= factor("EF_A"),Variable.Number= 9999, Variable.Name= factor("YEAR"))
EFA_var <- variables[variables$Table == "EF_A", 3]

# Merge and and only use the columns in our variable list

EFA <- lapply(EFA, function(x) x[(names(x)) %in% EFA_var])
EFA_df <- do.call("smartbind", EFA)

EFA_df<-EFA_df[, -3]

EFA_df$LINE <- as.character(EFA_df$LINE)

EFA_df <- EFA_df %>% filter(LINE == "8" | LINE == "22"| LINE == "29")

EFA_df$LINE[EFA_df$LINE == "8"] <- "Undergraduates"
EFA_df$LINE[EFA_df$LINE == "22"] <- "Undergraduates"
EFA_df$LINE[EFA_df$LINE == "29"] <- "Total"




EFA_df <- EFA_df %>% group_by(UNITID, LINE, YEAR) %>% summarize(EFTOTLM = sum(EFTOTLM, na.rm=TRUE), EFTOTLT = sum(EFTOTLT, na.rm = TRUE), EFHISPT = sum(EFHISPT, na.rm = TRUE))

EFA_df$YEAR <- ifelse(nchar(EFA_df$YEAR) != 4, gsub(" ", "", paste("19", EFA_df$YEAR), fixed = TRUE), EFA_df$YEAR)


library(tidyr)
EFA_df <- EFA_df %>% gather(4:6, key = "Type", value = "Count")
EFA_df <- EFA_df %>% spread(key = LINE, value = Count)
EFA_df$Graduate <- EFA_df$Total - EFA_df$Undergraduates
EFA_df <- EFA_df %>% gather(Total, Undergraduates, Graduate, key = "LINE", value = "Count")
EFA_df <- EFA_df %>% spread(Type, Count)


```

* Fall Enrollment - Major field of study, race/ethnicity, gender, attendance status, and level of student

WE WILL NOT USE THIS - MAJOR DOES NOT INCLUDE THEOLOGY?RELIGIOUS STUDIES

The code below acquires data for the major field of study enrollment data. The data is available for even years between 2000-2016.

*I can easily write code to acquire the data for year before 2000 if needed.*

```{r, warning= FALSE, message=FALSE}
survey_file <- "ef"
capital_survey_file <- "EF"
extra <- "cp"
capital_extra <- "CP"

for (year in 2000:2016){
  try(
  assign(paste(survey_file, year, extra, sep = ""), 
         getIPEDSData(year, survey_file, capital_survey_file, extra, capital_extra)))
}

#2006 
#Grand total
colnames(ef2006cp)[colnames(ef2006cp)=="EFRACE24"] <- "EFTOTLT"

# total men
colnames(ef2006cp)[colnames(ef2006cp)=="EFRACE15"] <- "EFTOTLM"

# total hispanic
colnames(ef2006cp)[colnames(ef2006cp)=="EFRACE21"] <- "EFHISPT"

#2004 
#Grand total

colnames(ef2004cp)[colnames(ef2004cp) == "EFRACE24"] <- "EFTOTLT"

#total men

colnames(ef2004cp)[colnames(ef2004cp) == "EFRACE15"] <- "EFTOTLM"

#total hispanic

colnames(ef2004cp)[colnames(ef2004cp)=="EFRACE21"] <- "EFHISPT"

#2002
colnames(ef2002cp)[colnames(ef2002cp) == "EFRACE24"] <- "EFTOTLT"

#total men

colnames(ef2002cp)[colnames(ef2002cp) == "EFRACE15"] <- "EFTOTLM"

#total hispanic

colnames(ef2002cp)[colnames(ef2002cp)=="EFRACE21"] <- "EFHISPT"

#2000

#2000 Total Men
colnames(ef2000cp)[colnames(ef2000cp)=="EFRACE15"] <- "EFTOTLM"
#2000 Total Women
colnames(ef2000cp)[colnames(ef2000cp)=="EFRACE16"] <- "EFTOTLW"
#total men and women
ef2000cp$EFTOTLT <- ef2000cp$EFTOTLM + ef2000cp$EFTOTLW

#2001 total Hispanic 
colnames(ef2000cp)[colnames(ef2000cp)=="EFRACE09"] <- "EFHISPM"
colnames(ef2000cp)[colnames(ef2000cp)=="EFRACE10"] <- "EFHISPW"

#total hispanic men and women
ef2000cp$EFHISPT <- ef2000cp$EFHISPW + ef2000cp$EFHISPM
```

```{r}
EFC <- mget(ls(pattern = "^ef\\d{4}cp$"))

levels(variables$Variable.Name) <- c(levels(variables$Variable.Name),"YEAR") 
variables[nrow(variables) + 1,] = list(Table= factor("EF_CP"),Variable.Number= 9999, Variable.Name= factor("YEAR"))
EFC_var <- variables[variables$Table == "EF_CP", 3]

# Merge and and only use the columns in our variable list
EFC <- lapply(EFC, function(x) x[(names(x)) %in% EFC_var])
EFC_df <- do.call("smartbind", EFC)
```



# Institutional Characteristics - Educational offerings, organization, services and athletic associations

The code below acquires the institutional characteristics variables of interest, specifically religious affiliation for 2000-2017. *The data is available from 1980 if needed - I can easily write code to include earlier years if needed*.
```{r, warning = FALSE, message = FALSE}
survey_file <- "ic"
capital_survey_file <- "IC"
extra <- ""
capital_extra <- ""

#2000-2012. Available from 1980
for (year in 2000:2017){
    assign(paste(survey_file, year, extra, sep = ""), 
           getIPEDSData(year, survey_file, capital_survey_file, extra, capital_extra))
}

```

Then, we made one big data frame.
```{r, warning = FALSE, message = FALSE}
IC <- mget(ls(pattern = "^ic\\d{4}$"))

levels(variables$Variable.Name) <- c(levels(variables$Variable.Name),"YEAR") 
variables[nrow(variables) + 1,] = list(Table= factor("IC"),Variable.Number= 9999, Variable.Name= factor("YEAR"))
IC_var <- variables[variables$Table == "IC", 3]

# Merge and and only use the columns in our variable list
IC <- lapply(IC, function(x) x[(names(x)) %in% IC_var])
IC_df <- do.call("smartbind", IC)
```

```{r}
DF <- merge(IC_df, HD_df, by = c("UNITID", "YEAR"), all.x = TRUE, all.y = TRUE)
DF <- merge(DF, EFA_df, by.x = c("UNITID", "YEAR"), by.y = c("UNITID", "YEAR"), all.x = TRUE, all.y = TRUE)


write.csv(DF, "data.csv")
```





