---
title: "model-building"
author: "Sarah McDonald"
date: "February 1, 2020"
output:
  html_document: default
  pdf_document: default
  word_document: default
---

```{r}

library(data.table)
library(knitr)
library(dplyr)
library(kableExtra)
library(jtools)
library(huxtable)
library(ggplot2)

data <- read.csv("C:/Users/Sarah McDonald/Documents/honorsthesis/manipulated_data/cath_inst.csv")

#remove 1990 and 1991
data <- data[data$YEAR!= 1990 & data$YEAR != 1991 & data$YEAR != 1992, ]



```

### Coefficient, P-Value, R-Squared Summary table
```{r}
summary_table <- data.table(`Variable` = c("Total Enrollment Publicity Coefficient", "Total Enrollment P-Value", "Total Enrollment R-squared", "Undergraduate Enrollment Publicity Coefficient", "Undergraduate Enrollment P-Value", "Undergraduate Enrollment R-squared", "Graduate Enrollment Publicity Coefficient", "Graduate Enrollment P-Value", "Graduate Enrollment R-squared"), `Nexis-Lexus: Linear` = numeric(length = 9), `Nexis-Lexus: Log(X+1)` = numeric(length = 9), `Nexis-Lexus: Square Root` = numeric(length = 9), `Bishop Accountability: Linear` = numeric(length = 9), `Bishop Accountability: Log(X+1)` = numeric(length = 9), `Bishop Accountability: Square Root` = numeric(length = 9))


```

# Bishop Accountability 

### TOTAL ENROLLMENT 
```{r}
enroll_total <- lm(EFTOTLT ~   bishop_pub_sum_3  + Percent.Catholic + his_prop + pop_dens + income_per_capita, data=data[data$LINE == "Total", ]) 


summary_table[summary_table$Variable  == "Total Enrollment Publicity Coefficient","Bishop Accountability: Linear"] <- summary(enroll_total)$coefficients[2, 1]
summary_table[summary_table$Variable  == "Total Enrollment P-Value","Bishop Accountability: Linear"] <- summary(enroll_total)$coefficients[2, 4]
summary_table[summary_table$Variable  == "Total Enrollment R-squared","Bishop Accountability: Linear"] <-summary(enroll_total)$r.squared
```

### GRADUATE ENROLLMENT 
```{r}
enroll_grad <- lm(EFTOTLT ~   bishop_pub_sum_3 + Percent.Catholic + his_prop + pop_dens + income_per_capita, data=data[data$LINE == "Graduate", ]) 

summary(enroll_grad)

summary_table[summary_table$Variable  == "Graduate Enrollment Publicity Coefficient","Bishop Accountability: Linear"] <- summary(enroll_grad)$coefficients[2, 1]
summary_table[summary_table$Variable  == "Graduate Enrollment P-Value","Bishop Accountability: Linear"] <- summary(enroll_grad)$coefficients[2, 4]
summary_table[summary_table$Variable  == "Graduate Enrollment R-squared","Bishop Accountability: Linear"] <-summary(enroll_grad)$r.squared





```

### UNDERGRADUATE ENROLLMENT 
```{r}
enroll_undergrad <- lm(EFTOTLT ~ bishop_pub_sum_3   + Percent.Catholic + his_prop + pop_dens + income_per_capita, data=data[data$LINE == "Undergraduates", ]) 

summary(enroll_undergrad)



summary_table[summary_table$Variable  == "Undergraduate Enrollment Publicity Coefficient","Bishop Accountability: Linear"] <- summary(enroll_undergrad)$coefficients[2, 1]
summary_table[summary_table$Variable  == "Undergraduate Enrollment P-Value","Bishop Accountability: Linear"] <- summary(enroll_undergrad)$coefficients[2, 4]
summary_table[summary_table$Variable  == "Undergraduate Enrollment R-squared","Bishop Accountability: Linear"] <-summary(enroll_undergrad)$r.squared


```


###TOTAL ENROLLMENT - Square Root Transformation
```{r}
enroll_total_sqrt <- lm(sqrt(EFTOTLT) ~   sqrt(bishop_pub_sum_3) + Percent.Catholic + his_prop + pop_dens + income_per_capita, data=data[data$LINE == "Undergraduates", ]) 

summary(enroll_total_sqrt)

bishop_total_sqrt <- summ(enroll_total_sqrt)

summary_table[summary_table$Variable  == "Total Enrollment Publicity Coefficient","Bishop Accountability: Square Root"] <- summary(enroll_total_sqrt)$coefficients[2, 1]
summary_table[summary_table$Variable  == "Total Enrollment P-Value","Bishop Accountability: Square Root"] <- summary(enroll_total_sqrt)$coefficients[2, 4]
summary_table[summary_table$Variable  == "Total Enrollment R-squared","Bishop Accountability: Square Root"] <-summary(enroll_total_sqrt)$r.squared


```

### GRADUATE ENROLLMENT - Square Root Transformation
```{r}
enroll_grad_sqrt <- lm(sqrt(EFTOTLT) ~   sqrt(bishop_pub_sum_3) + Percent.Catholic + his_prop + pop_dens + income_per_capita, data=data[data$LINE == "Graduate", ]) 

summary(enroll_grad_sqrt)

bishop_grad_sqrt <- summ(enroll_grad_sqrt)


summary_table[summary_table$Variable  == "Graduate Enrollment Publicity Coefficient","Bishop Accountability: Square Root"] <- summary(enroll_grad_sqrt)$coefficients[2, 1]
summary_table[summary_table$Variable  == "Graduate Enrollment P-Value","Bishop Accountability: Square Root"] <- summary(enroll_grad_sqrt)$coefficients[2, 4]
summary_table[summary_table$Variable  == "Graduate Enrollment R-squared","Bishop Accountability: Square Root"] <-summary(enroll_grad_sqrt)$r.squared
```

### UNDERGRADUATE ENROLLMENT - Square Root Transformation
```{r}
enroll_undergrad_sqrt <- lm(sqrt(EFTOTLT) ~   sqrt(bishop_pub_sum_3)  + Percent.Catholic + his_prop + pop_dens + income_per_capita, data=data[data$LINE == "Undergraduates", ]) 

summary(enroll_undergrad_sqrt)

bishop_undergrad_sqrt <- summ(enroll_undergrad_sqrt)


summary_table[summary_table$Variable  == "Undergraduate Enrollment Publicity Coefficient","Bishop Accountability: Square Root"] <- summary(enroll_undergrad_sqrt)$coefficients[2, 1]
summary_table[summary_table$Variable  == "Undergraduate Enrollment P-Value","Bishop Accountability: Square Root"] <- summary(enroll_undergrad_sqrt)$coefficients[2, 4]
summary_table[summary_table$Variable  == "Undergraduate Enrollment R-squared","Bishop Accountability: Square Root"] <-summary(enroll_undergrad_sqrt)$r.squared
```



###TOTAL ENROLLMENT - Log+1 Transformation
```{r}
enroll_total_log <- lm(log(EFTOTLT+1) ~   log(bishop_pub_sum_3+1) + Percent.Catholic + his_prop + pop_dens + income_per_capita, data=data[data$LINE == "Undergraduates", ]) 

summary(enroll_total_log)

summary_table[summary_table$Variable  == "Total Enrollment Publicity Coefficient","Bishop Accountability: Log(X+1)"] <- summary(enroll_total_log)$coefficients[2, 1]
summary_table[summary_table$Variable  == "Total Enrollment P-Value","Bishop Accountability: Log(X+1)"] <- summary(enroll_total_log)$coefficients[2, 4]
summary_table[summary_table$Variable  == "Total Enrollment R-squared","Bishop Accountability: Log(X+1)"] <-summary(enroll_total_log)$r.squared
```

### GRADUATE ENROLLMENT - Log+1 Transformation
```{r}
enroll_grad_log <- lm(log(EFTOTLT+1) ~   log(bishop_pub_sum_3+1)  + Percent.Catholic + his_prop + pop_dens + income_per_capita, data=data[data$LINE == "Graduate", ]) 

summary(enroll_grad_log)


summary_table[summary_table$Variable  == "Graduate Enrollment Publicity Coefficient","Bishop Accountability: Log(X+1)"] <- summary(enroll_grad_log)$coefficients[2, 1]
summary_table[summary_table$Variable  == "Graduate Enrollment P-Value","Bishop Accountability: Log(X+1)"] <- summary(enroll_grad_log)$coefficients[2, 4]
summary_table[summary_table$Variable  == "Graduate Enrollment R-squared","Bishop Accountability: Log(X+1)"] <-summary(enroll_grad_log)$r.squared
```

### UNDERGRADUATE ENROLLMENT - Log+1 Transformation
```{r}
enroll_undergrad_log <- lm(log(EFTOTLT+1) ~   log(bishop_pub_sum_3+1)   + Percent.Catholic + his_prop + pop_dens + income_per_capita, data=data[data$LINE == "Undergraduates", ]) 

summary(enroll_undergrad_log)

summary_table[summary_table$Variable  == "Undergraduate Enrollment Publicity Coefficient","Bishop Accountability: Log(X+1)"] <- summary(enroll_undergrad_log)$coefficients[2, 1]
summary_table[summary_table$Variable  == "Undergraduate Enrollment P-Value","Bishop Accountability: Log(X+1)"] <- summary(enroll_undergrad_log)$coefficients[2, 4]
summary_table[summary_table$Variable  == "Undergraduate Enrollment R-squared","Bishop Accountability: Log(X+1)"] <-summary(enroll_undergrad_log)$r.squared
```



# Nexis-Lexus

### TOTAL ENROLLMENT 
```{r}
nexis_enroll_total <- lm(EFTOTLT ~   nexis_sum_3  + Percent.Catholic + his_prop + pop_dens + income_per_capita, data=data[data$LINE == "Total", ]) 

summary(nexis_enroll_total)

summary_table[summary_table$Variable  == "Total Enrollment Publicity Coefficient","Nexis-Lexus: Linear"] <- summary(nexis_enroll_total)$coefficients[2, 1]
summary_table[summary_table$Variable  == "Total Enrollment P-Value","Nexis-Lexus: Linear"] <- summary(nexis_enroll_total)$coefficients[2, 4]
summary_table[summary_table$Variable  == "Total Enrollment R-squared","Nexis-Lexus: Linear"] <-summary(nexis_enroll_total)$r.squared
```

### GRADUATE ENROLLMENT
```{r}
nexis_enroll_grad <- lm(EFTOTLT ~   nexis_sum_3 + Percent.Catholic + his_prop + pop_dens + income_per_capita, data=data[data$LINE == "Graduate", ]) 

summary(nexis_enroll_grad)

summary_table[summary_table$Variable  == "Graduate Enrollment Publicity Coefficient","Nexis-Lexus: Linear"] <- summary(nexis_enroll_grad)$coefficients[2, 1]
summary_table[summary_table$Variable  == "Graduate Enrollment P-Value","Nexis-Lexus: Linear"] <- summary(nexis_enroll_grad)$coefficients[2, 4]
summary_table[summary_table$Variable  == "Graduate Enrollment R-squared","Nexis-Lexus: Linear"] <-summary(nexis_enroll_grad)$r.squared
```

### UNDERGRADUATE ENROLLMENT 
```{r}
nexis_enroll_undergrad <- lm(EFTOTLT ~ nexis_sum_3   + Percent.Catholic + his_prop + pop_dens + income_per_capita, data=data[data$LINE == "Undergraduates", ]) 

summary(nexis_enroll_undergrad)

summary_table[summary_table$Variable  == "Undergraduate Enrollment Publicity Coefficient","Nexis-Lexus: Linear"] <- summary(nexis_enroll_undergrad)$coefficients[2, 1]
summary_table[summary_table$Variable  == "Undergraduate Enrollment P-Value","Nexis-Lexus: Linear"] <- summary(nexis_enroll_undergrad)$coefficients[2, 4]
summary_table[summary_table$Variable  == "Undergraduate Enrollment R-squared","Nexis-Lexus: Linear"] <-summary(nexis_enroll_undergrad)$r.squared
```



###TOTAL ENROLLMENT - Square Root Transformation
```{r}
nexis_enroll_total_sqrt <- lm(sqrt(EFTOTLT) ~   sqrt(nexis_sum_3) + Percent.Catholic + his_prop + pop_dens + income_per_capita, data=data[data$LINE == "Undergraduates", ]) 

summary(nexis_enroll_total_sqrt)

nexis_total_sqrt <- summ(nexis_enroll_total_sqrt)

summary_table[summary_table$Variable  == "Total Enrollment Publicity Coefficient","Nexis-Lexus: Square Root"] <- summary(nexis_enroll_total_sqrt)$coefficients[2, 1]
summary_table[summary_table$Variable  == "Total Enrollment P-Value","Nexis-Lexus: Square Root"] <- summary(nexis_enroll_total_sqrt)$coefficients[2, 4]
summary_table[summary_table$Variable  == "Total Enrollment R-squared","Nexis-Lexus: Square Root"] <-summary(nexis_enroll_total_sqrt)$r.squared
```

### GRADUATE ENROLLMENT - Square Root Transformation
```{r}
nexis_enroll_grad_sqrt <- lm(sqrt(EFTOTLT) ~   sqrt(nexis_sum_3) + Percent.Catholic + his_prop + pop_dens + income_per_capita, data=data[data$LINE == "Graduate", ]) 

summary(nexis_enroll_grad_sqrt)

nexis_grad_sqrt <- summ(nexis_enroll_grad_sqrt)


summary_table[summary_table$Variable  == "Graduate Enrollment Publicity Coefficient","Nexis-Lexus: Square Root"] <- summary(nexis_enroll_grad_sqrt)$coefficients[2, 1]
summary_table[summary_table$Variable  == "Graduate Enrollment P-Value","Nexis-Lexus: Square Root"] <- summary(nexis_enroll_grad_sqrt)$coefficients[2, 4]
summary_table[summary_table$Variable  == "Graduate Enrollment R-squared","Nexis-Lexus: Square Root"] <-summary(nexis_enroll_grad_sqrt)$r.squared
```

### UNDERGRADUATE ENROLLMENT - Square Root Transformation
```{r}
nexis_enroll_undergrad_sqrt <- lm(sqrt(EFTOTLT) ~   sqrt(nexis_sum_3)  + Percent.Catholic + his_prop + pop_dens + income_per_capita, data=data[data$LINE == "Undergraduates", ]) 

summary(nexis_enroll_undergrad_sqrt)

nexis_undergrad_sqrt <- summ(nexis_enroll_undergrad_sqrt)

summary_table[summary_table$Variable  == "Undergraduate Enrollment Publicity Coefficient","Nexis-Lexus: Square Root"] <- summary(nexis_enroll_undergrad_sqrt)$coefficients[2, 1]
summary_table[summary_table$Variable  == "Undergraduate Enrollment P-Value","Nexis-Lexus: Square Root"] <- summary(nexis_enroll_undergrad_sqrt)$coefficients[2, 4]
summary_table[summary_table$Variable  == "Undergraduate Enrollment R-squared","Nexis-Lexus: Square Root"] <-summary(nexis_enroll_undergrad_sqrt)$r.squared
```


###TOTAL ENROLLMENT - Log + 1 Transformation
```{r}
nexis_enroll_total_log <- lm(log(EFTOTLT+1) ~   log(nexis_sum_3+1) + Percent.Catholic + his_prop + pop_dens + income_per_capita, data=data[data$LINE == "Undergraduates", ]) 

summary(nexis_enroll_total_log)

summary_table[summary_table$Variable  == "Total Enrollment Publicity Coefficient","Nexis-Lexus: Log(X+1)"] <- summary(nexis_enroll_total_log)$coefficients[2, 1]
summary_table[summary_table$Variable  == "Total Enrollment P-Value","Nexis-Lexus: Log(X+1)"] <- summary(nexis_enroll_total_log)$coefficients[2, 4]
summary_table[summary_table$Variable  == "Total Enrollment R-squared","Nexis-Lexus: Log(X+1)"] <-summary(nexis_enroll_total_log)$r.squared
```

### GRADUATE ENROLLMENT - Log + 1 Transformation
```{r}
nexis_enroll_grad_log <- lm(log(EFTOTLT+1) ~   log(nexis_sum_3+1) + Percent.Catholic + his_prop + pop_dens + income_per_capita, data=data[data$LINE == "Graduate", ]) 

summary(nexis_enroll_grad_log)

summary_table[summary_table$Variable  == "Graduate Enrollment Publicity Coefficient","Nexis-Lexus: Log(X+1)"] <- summary(nexis_enroll_grad_log)$coefficients[2, 1]
summary_table[summary_table$Variable  == "Graduate Enrollment P-Value","Nexis-Lexus: Log(X+1)"] <- summary(nexis_enroll_grad_log)$coefficients[2, 4]
summary_table[summary_table$Variable  == "Graduate Enrollment R-squared","Nexis-Lexus: Log(X+1)"] <-summary(nexis_enroll_grad_log)$r.squared
```

### UNDERGRADUATE ENROLLMENT - Log + 1 Transformation
```{r}
nexis_enroll_undergrad_log <- lm(log(EFTOTLT+1) ~   log(nexis_sum_3+1) + Percent.Catholic + his_prop + pop_dens + income_per_capita, data=data[data$LINE == "Undergraduates", ]) 

summary(nexis_enroll_undergrad_log)

summary_table[summary_table$Variable  == "Undergraduate Enrollment Publicity Coefficient","Nexis-Lexus: Log(X+1)"] <- summary(nexis_enroll_undergrad_log)$coefficients[2, 1]
summary_table[summary_table$Variable  == "Undergraduate Enrollment P-Value","Nexis-Lexus: Log(X+1)"] <- summary(nexis_enroll_undergrad_log)$coefficients[2, 4]
summary_table[summary_table$Variable  == "Undergraduate Enrollment R-squared","Nexis-Lexus: Log(X+1)"] <-summary(nexis_enroll_undergrad_log)$r.squared
```



# Percent Change Table
```{r}
change_table <- data.table(`Variable` = c("Total Enrollment % Change Publicity Coefficient", "Total Enrollment % Change P-Value", "Total Enrollment % Change R-squared", "Undergraduate Enrollment % Change Publicity Coefficient", "Undergraduate Enrollment % Change P-Value", "Undergraduate Enrollment % Change R-squared", "Graduate Enrollment % Change Publicity Coefficient", "Graduate Enrollment % Change P-Value", "Graduate Enrollment % Change R-squared"), `Nexis-Lexis: Linear` = numeric(length = 9),  `Bishop Accountability: Linear` = numeric(length = 9))


```



# Bishop Accountability Percent change

### TOTAL ENROLLMENT 
```{r}
bishop_total_change <- lm(EFTOTLT_3_change ~   bishop_pub_sum_3  + Percent.Catholic + his_prop + pop_dens + income_per_capita, data=data[data$LINE == "Total", ]) 

summary(bishop_total_change)

change_table[change_table$Variable  == "Total Enrollment % Change Publicity Coefficient","Bishop Accountability: Linear"] <- summary(bishop_total_change)$coefficients[2, 1]
change_table[change_table$Variable  == "Total Enrollment % Change P-Value","Bishop Accountability: Linear"] <- summary(bishop_total_change)$coefficients[2, 4]
change_table[change_table$Variable  == "Total Enrollment % Change R-squared","Bishop Accountability: Linear"] <-summary(bishop_total_change)$r.squared

```

### GRADUATE ENROLLMENT 
```{r}
bishop_grad_change <- lm(EFTOTLT_3_change ~   bishop_pub_sum_3 + Percent.Catholic + his_prop + pop_dens + income_per_capita, data=data[data$LINE == "Graduate", ]) 

summary(bishop_grad_change)

change_table[change_table$Variable  == "Graduate Enrollment % Change Publicity Coefficient","Bishop Accountability: Linear"] <- summary(bishop_grad_change)$coefficients[2, 1]
change_table[change_table$Variable  == "Graduate Enrollment % Change P-Value","Bishop Accountability: Linear"] <- summary(bishop_grad_change)$coefficients[2, 4]
change_table[change_table$Variable  == "Graduate Enrollment % Change R-squared","Bishop Accountability: Linear"] <-summary(bishop_grad_change)$r.squared
```

### UNDERGRADUATE ENROLLMENT 
```{r}
bishop_undergrad_change <- lm(EFTOTLT_3_change ~ bishop_pub_sum_3   + Percent.Catholic + his_prop + pop_dens + income_per_capita, data=data[data$LINE == "Undergraduates", ]) 

summary(bishop_undergrad_change)

change_table[change_table$Variable  == "Undergraduate Enrollment % Change Publicity Coefficient","Bishop Accountability: Linear"] <- summary(bishop_undergrad_change)$coefficients[2, 1]
change_table[change_table$Variable  == "Undergraduate Enrollment % Change P-Value","Bishop Accountability: Linear"] <- summary(bishop_undergrad_change)$coefficients[2, 4]
change_table[change_table$Variable  == "Undergraduate Enrollment % Change R-squared","Bishop Accountability: Linear"] <-summary(bishop_undergrad_change)$r.squared

```


# Percent Change, Nexis

### TOTAL ENROLLMENT 
```{r}
nexis_total_change <- lm(EFTOTLT_3_change ~   nexis_sum_3  + Percent.Catholic + his_prop + pop_dens + income_per_capita, data=data[data$LINE == "Total", ]) 

summary(nexis_total_change)

change_table[change_table$Variable  == "Total Enrollment % Change Publicity Coefficient","Nexis-Lexis: Linear"] <- summary(nexis_total_change)$coefficients[2, 1]
change_table[change_table$Variable  == "Total Enrollment % Change P-Value","Nexis-Lexis: Linear"] <- summary(nexis_total_change)$coefficients[2, 4]
change_table[change_table$Variable  == "Total Enrollment % Change R-squared","Nexis-Lexis: Linear"] <-summary(nexis_total_change)$r.squared

```

### GRADUATE ENROLLMENT 
```{r}
nexis_grad_change <- lm(EFTOTLT_3_change ~   nexis_sum_3 + Percent.Catholic + his_prop + pop_dens + income_per_capita, data=data[data$LINE == "Graduate", ]) 

summary(nexis_grad_change)

change_table[change_table$Variable  == "Graduate Enrollment % Change Publicity Coefficient","Nexis-Lexis: Linear"] <- summary(nexis_grad_change)$coefficients[2, 1]
change_table[change_table$Variable  == "Graduate Enrollment % Change P-Value","Nexis-Lexis: Linear"] <- summary(nexis_grad_change)$coefficients[2, 4]
change_table[change_table$Variable  == "Graduate Enrollment % Change R-squared","Nexis-Lexis: Linear"] <-summary(nexis_grad_change)$r.squared

```

### UNDERGRADUATE ENROLLMENT 
```{r}
nexis_undergrad_change <- lm(EFTOTLT_3_change ~ nexis_sum_3   + Percent.Catholic + his_prop + pop_dens + income_per_capita, data=data[data$LINE == "Undergraduates", ]) 

summary(nexis_undergrad_change)

change_table[change_table$Variable  == "Undergraduate Enrollment % Change Publicity Coefficient","Nexis-Lexis: Linear"] <- summary(nexis_undergrad_change)$coefficients[2, 1]
change_table[change_table$Variable  == "Undergraduate Enrollment % Change P-Value","Nexis-Lexis: Linear"] <- summary(nexis_undergrad_change)$coefficients[2, 4]
change_table[change_table$Variable  == "Undergraduate Enrollment % Change R-squared","Nexis-Lexis: Linear"] <-summary(nexis_undergrad_change)$r.squared


```


```{r}
kable(summary_table) %>%
  row_spec(1:3, bold = T, color = "white", background = "#57A773") %>%
  row_spec(4:6, bold = T, color = "white", background = "#A7A5C6") %>%
  row_spec(7:9, bold = T, color = "white", background = "#EDCF8E") 

kable(change_table) %>%
  row_spec(1:3, bold = T, color = "white", background = "#57A773") %>%
  row_spec(4:6, bold = T, color = "white", background = "#A7A5C6") %>%
  row_spec(7:9, bold = T, color = "white", background = "#EDCF8E") 


```


```{r}
export_summs(bishop_total_sqrt, bishop_grad_sqrt, bishop_undergrad_sqrt, nexis_total_sqrt, nexis_grad_sqrt, nexis_undergrad_sqrt, model.names = c("Bishop Accountability: Total", "Bishop Accountability: Graduate", "Bishop Accountability: Undergraduate", "Nexis-Lexus: Total", "Nexis-Lexus: Graduate", "Nexis-Lexus: Undergraduate"))



export_summs(bishop_total_change, bishop_grad_change, bishop_undergrad_change, nexis_total_change, nexis_grad_change, nexis_undergrad_change, model.names = c("Bishop Accountability: Total % Change", "Bishop Accountability: Graduate % Change", "Bishop Accountability: Undergraduate % Change", "Nexis-Lexus: Total % Change", "Nexis-Lexus: Graduate % Change", "Nexis-Lexus: Undergraduate % Change"))
```





It is interesting that these are not the same top five
```{r}

bishop_pub <- read.csv("C:/Users/Sarah McDonald/Documents/honorsthesis/manipulated_data/bishop_pub.csv")
#bishop_pub <- 
  
  bishop_pub %>% filter(public < 2018 & 1989 < public ) %>% group_by(Diocese) %>% summarise(total = sum(bishop_pub)) %>% arrange(desc(total)) %>% top_n(10, total)
  
  bishop_figure <-bishop_pub %>% filter(public < 2018 & 1989< public) %>% group_by(public) %>% summarize(total = n())

  ggplot(bishop_figure, aes(x=public, y=total)) +
  geom_line() +
  labs(title = "Bishop Publicity: Count of Negative Publicity for Priests Over Time")
  
  
nexus <- read.csv("C:/Users/Sarah McDonald/Documents/honorsthesis/manipulated_data/nexus.csv")
#bishop_pub <- 
  
  nexus %>% filter(year < 2018 & 1989 < year ) %>% group_by(dioceses) %>% summarise(total = sum(nexus)) %>% arrange(desc(total)) %>% top_n(10, total)
  
  
  
  nexus_figure <- nexus %>% filter(year < 2018 & 1989 < year) %>% group_by(year) %>% summarise(total = sum(nexus))

  ggplot(nexus_figure, aes(x=year, y=total)) +
  geom_line() +
  labs(title = "NexisUni: Count of Negative Newswires Over Time")
```


Phillly, Boston , new York,  LA
```{r}





boston <- data %>% filter(Dioceses == "Boston"| Dioceses == "New York" | Dioceses == "Philadelphia"| Dioceses == "Los Angeles")


boston_lm  <- lm(EFTOTLT_3_change ~   bishop_pub_sum_3 + Percent.Catholic + his_prop + pop_dens + income_per_capita, data=boston) 

summary(boston_lm)

ggplot(boston, aes(x=YEAR, y=bishop_pub)) +
  geom_line() 

ggplot(boston, aes(x=YEAR, y=nexis)) +
  geom_line() 

ggplot(boston[boston$LINE == "Total", ], aes(x=YEAR, y=EFTOTLT, color = INSTNM)) +
  geom_line() 

```












# Seminaries

```{r}
data$CARNEGIE <- as.character(data$CARNEGIE)

seminary.ids <- data[data$CARNEGIE == "Theological seminaries and other specialized faith-related institutions" & is.na(data$CARNEGIE)==FALSE, "UNITID"]

seminary.ids <- unique(seminary.ids)


seminary <- data[data$UNITID %in% seminary.ids, ]

seminary_bishop <- lm(EFTOTLT ~   bishop_pub_sum_3 + Percent.Catholic + his_prop + pop_dens + income_per_capita, data=seminary) 

summary(seminary_bishop)

seminary_nexis <- lm(EFTOTLT ~  nexis_sum_3 + Percent.Catholic + his_prop + pop_dens + income_per_capita, data=seminary) 

summary(seminary_nexis)




```












# Other Things I Have Tried

```{r}
library(ggplot2)

ggplot(data[data$LINE == "Total", ], aes(x= nexis_sum_3, y = EFTOTLT))+
  geom_point() +
  theme_classic() +
  labs(title = "Nexis Publicity VS Total Enrollment", x = "Nexis Publicity", y = "Total Enrollment")


ggplot(data[data$LINE == "Total", ], aes(x= bishop_pub_sum_3, y = EFTOTLT))+
  geom_point()+
  theme_classic()+
  labs(title = "Bishop Accountability Publicity VS Total Enrollment", x = "Bishop Accountability Publicity", y = "Total Enrollment")


```

```{r}

ggplot(data[data$LINE == "Total", ], aes(x= log(nexis_sum_3+0.0000000001), y = log(EFTOTLT+0.0000000001)))+
  geom_point() +
  theme_classic() +
  labs(title = "Nexis Publicity VS Total Enrollment \n(Log (X+0.0000000001) Transformation)", x = "Nexis Publicity", y = "Total Enrollment")

ggplot(data[data$LINE == "Total", ], aes(x= log(bishop_pub_sum_3+0.0000000001), y = log(EFTOTLT+0.0000000001)))+
  geom_point()+
  theme_classic()+
  labs(title = "Bishop Accountability Publicity VS Total Enrollment \n(Log (X+0.0000000001) Transformation)", x = "Bishop Accountability Publicity", y = "Total Enrollment")
```


```{r}
ggplot(data[data$LINE == "Total", ], aes(x= sqrt(nexis_sum_3), y = sqrt(EFTOTLT)))+
  geom_point() +
  theme_classic() +
  labs(title = "Nexis Publicity VS Total Enrollment \n(SQRT Transformation)", x = "Nexis Publicity", y = "Total Enrollment")

ggplot(data[data$LINE == "Total", ], aes(x= sqrt(bishop_pub_sum_3), y = sqrt(EFTOTLT)))+
  geom_point()+
  theme_classic()+
  labs(title = "Bishop Accountability Publicity VS Total Enrollment \n(SQRT Transformation)", x = "Bishop Accountability Publicity", y = "Total Enrollment")
```



zip model
```{r}


M1 <- glm(EFTOTLT ~ nexis_sum_3 + Percent.Catholic + his_prop + pop_dens + income_per_capita, data=data[data$LINE == "Total", ],
          family = 'poisson')

summary(M1)


E2 <- resid(M1, type = "pearson")
N  <- nrow(data)
p  <- length(coef(M1))   
sum(E2^2) / (N - p)
#overdispersion
```

```{r}
library(MASS)
M2 <- glm.nb(EFTOTLT ~ nexis_sum_3 + Percent.Catholic + his_prop + pop_dens + income_per_capita, data=data[data$LINE == "Total",])
summary(M2)

E2 <- resid(M2, type = "pearson")
N  <- nrow(data)
p  <- length(coef(M2)) + 1  # '+1' is for variance parameter in NB
sum(E2^2) / (N - p)
#underdispersion?

```

```{r}
#M3 <- zeroinfl(EFTOTLT ~ nexis_sum_3 + Percent.Catholic + his_prop + pop_dens + income_per_capita,
              # dist = 'poisson',
              # data=data[data$LINE == "Total",])

#summary(M3)
```

